---
title: "Class 7 R functions and packages"
author: "Claudia Medrano"
date: "10/23/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
source("http://tinyurl.com/rescale-R")
```

Let's try our rescale() function from last class
```{r}
rescale(1:10)
```

```{r}
rescale(c(3,10,NA,7))
```

# Test fail
rescale( c(1,10,"string") )

```{r}
rescale(c(3,10,NA))
```


stop("some message") to report on unexpected input type or other problem early in a function
rescale2(c(3,10,NA,"Barry")) would give an error

#write a function both_na()
We want to write a function, called both_na(),
that counts how many positions in two input
vectors, x and y, both have a missing value

```{r}
# from stackOverflow
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)

is.na(x)
is.na(y)
```

```{r}
which(is.na(x))
```

working snippet of code
```{r}
is.na(x) & is.na(y)
```

```{r}
both_na <- function(x,y) {
 sum(is.na(x) & is.na(y))
}
```

```{r}
both_na(x,y)
```

```{r}
x <- c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)
y3 <- c(1, NA, NA, NA, NA, NA, NA)

both_na(x,y3)
#error due to recycling
```

both_na2 <- function(x,y) {
 if(length(x) != length(y)) {
   stop("vector lengths are not the same")
 }
  sum(is.na(x) & is.na(y))
}

both_na2(x, y3)



Write a function grade() to determine an overall
grade from a vector of student homework
assignment scores dropping the lowest single
alignment score.
 student 1
c(100, 100, 100, 100, 100, 100, 100, 90)
 student 2
c(100, NA, 90, 90, 90, 90, 97, 80)

```{r}
s1 <- c(100, 100, 100, 100, 100, 100, 100, 90)
s2 <- c(100, NA, 90, 90, 90, 90, 97, 80)

#use which.min() from stack overflow
which.min(s2)
```
```{r}
#exclude it from average 
s1[which.min(s1)]
```

```{r}
#vector[-IND]
s1[-which.min(s1)]
s2[-which.min(s2)]
mean(s1[-which.min(s1)])
mean(s2[-which.min(s2)])
```

```{r}
mean(s2[-which.min(s2)], na.rm = TRUE)
```


```{r}
#student with the following scores would average 100
s3 <- c(100, NA, NA, NA, NA)
```

```{r}
x <- s3
sum(x[-which.min(x)], na.rm = TRUE) / (length(x) - 1)
```


