---
title: "class09"
author: "Claudia Medrano"
date: "10/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
wisc.df <- read.csv("WisconsinCancer.csv")
```

```{r}
head(wisc.df)
```

Here we examine data from `r nrow(wisc.df)` patient samples.

```{r}
x <- table(wisc.df$diagnosis)
x
```

In this data set we have `r x["B"]` non-cancer and `r x["M"]` cancer

We will not consider the columns `id`, `diagnosis`, or the last `X` column. We will take columns 3 to 32 only 

```{r}
# Convert the features of the data: wisc.data
wisc.data <- as.matrix(wisc.df[,3:32])
```

Q. How many variables/features in the data are suffixed with _mean?
```{r}
#use colnames
colnames(wisc.df)
```

```{r}
#use grep
grep("_mean", colnames(wisc.df), value = TRUE)
```

To find out how many there are I can call `length()`  on `grep()`
```{r}
#count amount of _mean
length(grep("_mean", colnames(wisc.df), value = TRUE))
```

Perform PCA on wisc.data

## Do we need to scale the data?
```{r}
round(apply(wisc.data, 2, sd),3)
```

Standard deviations vary greatly, so we will need to use `scale = TRUE`
```{r}
wisc.pr <- prcomp(wisc.data, scale=TRUE)
summary(wisc.pr)
```

plot PC1 vs PC2 and color by M/B cancer/non-cancer diagnosis
```{r}
#without color 
plot(wisc.pr$x[,1], wisc.pr$x[,2])
```

```{r}
#with color 
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=wisc.df$diagnosis, xlab="PC1", ylab="PC2")
```

Plot PC1 vs PC3
```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,3], col=wisc.df$diagnosis, xlab="PC1", ylab="PC3")
```

```{r}
x <- summary(wisc.pr)
x$importance[,"PC1"]
```
The first PC captures `r x$importance[2,"PC1"] * 100` of the original variance of the data set


How many principal components (PCs) are required to describe at least 70% of the original variance in the data?

```{r}
which(x$importance[3,] >0.7)[1]
```

##Hierarchical clustering of case data
```{r}
# Scale the wisc.data data: data.scaled
data.scaled <- scale(wisc.data)
```
```{r}
#euclidian dist
data.dist <- dist(data.scaled)
```

Create a hierarchical clustering model using complete linkage. Manually specify the method argument to hclust() and assign the results to wisc.hclust.

```{r}
wisc.hclust <- hclust(data.dist, method="complete")
```
Plot the clustering model
```{r}
plot(wisc.hclust)
```

Which height will the clustering model have 4 clusters?
```{r}
plot(wisc.hclust)
abline(h=19.5, col="red", lty=2 )
```

```{r}
wisc.hclust.clusters <- cutree(wisc.hclust, k=4)
```

use the table() function to compare the cluster membership to the actual diagnoses.

```{r}
diagnosis <- wisc.df$diagnosis
table(wisc.hclust.clusters, diagnosis)
```

##Clustering on PCA results

```{r}
wisc.pr.hclust <- hclust(dist(wisc.pr$x[,1:7]), method="ward.D2")
plot(wisc.pr.hclust)
```

```{r}
grps <- cutree(wisc.pr.hclust, h=70)
table(grps)
table(grps, diagnosis)
plot(wisc.pr$x[,1:2], col=grps)
```

```{r}
plot(wisc.pr$x[,1:2], col=diagnosis)
```

```{r}
g <- as.factor(grps)
levels(g)

g <- relevel(g,2)
levels(g)
plot(wisc.pr$x[,1:2], col=g)
```

We will use the predict() function that will take our PCA model from before and new cancer cell data and project that data onto our PCA space.

```{r}
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc
```
```{r}
plot(wisc.pr$x[,1:2], col=g)
points(npc[,1], npc[,2], col="blue", pch=16, cex=3)
text(npc[,1], npc[,2], c(1,2), col="white")
```
Which of these new patients should we prioritize for follow up based on your results?
2
