---
title: "class_08"
author: "Claudia Medrano"
date: "10/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Generate some example data for clustering
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))

plot(x)
```

Use the kmeans() function setting k to 2 and nstart=20


```{r}
k <-  kmeans(x, centers = 2, nstart = 20)

```
 
Inspect/print the results
```{r}
k
```

Q. How many points are in each cluster?
```{r}
k$size
```

Q. What ‘component’ of your result object details
 - cluster size?
 size
 - cluster assignment/membership?
```{r}
k$cluster
```
```{r}
table(k$cluster)
```
 
 - cluster center?
```{r}

```
 
 
Plot x colored by the kmeans cluster assignment and
 add cluster centers as blue points
 
```{r}
plot(x, col = k$cluster)
points(k$centers, col="blue", pch = 15)
```
 
 ##Hierarchical clustering in R
 Call:
hclust(d = dist_matrix)
Cluster method : complete
Distance : euclidean
Number of objects: 60 
```{r}
hc <- hclust(dist(x))
```
plot my results
```{r}
plot(hc)
#draws the red line at specified height
abline(h=6,col="red")
abline(h=4,col="blue")
cutree(hc, h=6)
```

```{r}
grps <- cutree(hc, h=4)
table(grps)
```

I can also cut the tree to yield a given 'k' groups/clusters
```{r}
grps <- cutree(hc, k=2)
plot(x, col=grps)
```

##Another cluster example
Here we have some more realistic 
```{r}
# Step 1. Generate some example data for clustering
x <- rbind(
 matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2),
 matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2),
 matrix(c(rnorm(50, mean = 1, sd = 0.3),
 rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")
# Step 2. Plot the data without clustering
plot(x)
# Step 3. Generate colors for known clusters
# (just so we can compare to hclust results)
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x, col=col)
```
 
 Use the dist(), hclust(), plot() and 
```{r}
grps <- hclust(dist(x))
plot(grps)
abline(h=2.4,col="blue")
abline(h=2, col="red")
```
 cutree() functions to return 2 and 3 clusters
```{r}
grps <- cutree(hc,h=1.8)
plot(grps,col=col)
```
 
Q. How does this compare to your known 'col' groups?


##Hands on PCA
```{r}
mydata <- read.csv("https://tinyurl.com/expression-CSV",
 row.names=1) 

head(mydata)
```

call prcomp on data set. we need to take the 't()' of our data
```{r}
pca <- prcomp(t(mydata), scale=TRUE) 
attributes(pca)
nrow(mydata)
ncol(mydata)
```
```{r}
pca$x
```

plot PCA1 vs PCA2
```{r}
plot(pca$x[,1], pca$x[,2])
```

```{r}
summary(pca)
```

```{r}
pca.var <- pca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)

barplot(pca.var.per, main="Scree Plot",
 xlab="Principal Component", ylab="Percent Variation")

```

```{r}
plot(pca$x[,1:2], col=c("red", "red", "red", "red", "red", "blue", "blue", "blue", "blue", "blue"))
```

##Perform a PCA on the UK foods dataset
```{r}
x <- read.csv(file = "data/UK_foods.csv")
```
Q1. How many rows and columns are in your new data frame named x? What R functions could you use to answer this questions?
```{r}
ncol(x)
nrow(x)
```

Preview the first 6 rows
```{r}
head(x)
```
```{r}
rownames(x) <- x[,1]
x <- x[,-1]
head(x)
```

alternative to fixing the columns 
```{r}
x <- read.csv(file = "data/UK_foods.csv", row.names = 1)
head(x)
```

Q2. Which approach to solving the ‘row-names problem’ mentioned above do you prefer and why? Is one approach more robust than another under certain circumstances?
set the row.names when loading the data table

```{r}
barplot(as.matrix(x), beside=T, col=rainbow(nrow(x)))
```

Q3: Changing what optional argument in the above barplot() function results in the following plot?


```{r}
barplot(as.matrix(x), beside=F, col=rainbow(nrow(x)))
```

Q5: Generating all pairwise plots may help somewhat. Can you make sense of the following code and resulting figure? What does it mean if a given point lies on the diagonal for a given plot?
```{r}
pairs(x, col=rainbow(10), pch=16)
```

Use PCA
```{r}
pca <- prcomp(t(x))
summary(pca)
```

plot
```{r}
plot(pca$x[,1:2])
```

```{r}
plot(pca$x[,1:2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1], pca$x[,2], colnames(x))
abline(h=0, col="gray", lt=2)
abline(v=0, col="gray", lt=2)
```

Lets focus on PC1 as it accounts for > 90% of variance 
```{r}
par(mar=c(10, 3, 0.35, 0))
barplot( pca$rotation[,1], las=2 )
```

