---
title: "Class 6 practice"
author: "Claudia Medrano"
date: "10/18/2019"
output: github_document
---

```{r}
read.table("test1.txt", sep = ",",  header = TRUE)
```

```{r}
read.table("test2.txt", sep = "$", header = TRUE)
```

```{r}
read.table("test3.txt")
```

```{r}
add <- function(x, y=1) {
 # Sum the input x and y
 x + y
}
```

```{r}
add( x = c(5,6), y = 100)
```


```{r}
add( c(5, 5, NA, 7) )
```






```{r}
min(c(5, 2, 7, 10))
```
```{r}
range(c(5,2,7,10))
```


```{r}
rescale <- function(x) {
 rng <-range(x)
 (x - rng[1]) / (rng[2] - rng[1])
}

```

```{r}
rescale(1:10)
```

```{r}
rescale(c(1,2,NA,3,10))
```

```{r}
x <- c(1,2,NA,5)
range(x)
```

flip the na.rm to be true (default is false)
```{r}
rescale2 <- function(x) {
 rng <-range(x, na.rm = TRUE)
 (x - rng[1]) / (rng[2] - rng[1])
}
```

```{r}
rescale2 <- function(x)
  rng <- range(x, na.rm = TRUE)

```

```{r}
rescale2(c(1,2,NA,3,10))
```

install the bio3d package with
```{r}
#install.packages("bio3d")
```

Exercise 1B

```{r}
#use library to load a package
library(bio3d)

#improving analysis code
library(bio3d)
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug

#make vector for entries
entries <- c("4AKE", "1AKE", "1E4Y")

#use read.pdb on each entry in entries vector
#for(x in entries) {
 # s_pdb <- read.pdb(x)
#}

s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")

#trim the lists in s_pdb and assign to new vector s_chain A
#for(s in s_pdb) {
 # s_chainA <- trim.pdb(s, chain="A", elety="CA")
#}

s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b

#
#for(chain in s_chainA) {
 # s.b_vector <- chain$atom$b
#}
  
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")

#plot 
#for(s.b in s.b_vector) {
#  plotb3(s.b, )
#}
```

Q1. What type of object is returned from the read.pdb() function?
 list 

Q2. What does the trim.pdb() function do?
makes the PDB object smaller by including only some subsets

Q3. What input parameter would turn off the marginal black and grey rectangles in the
plots and what do they represent in this case?
Q4. What would be a better plot to compare across the different proteins?
Q5. Which proteins are more similar to each other in their B-factor trends. How could
you quantify this? HINT: try the rbind(), dist() and hclust() functions together with a
resulting dendrogram plot. Look up the documentation to see what each of these
functions does.

